General constraints on magic distillation protocols from statistical mechanics

### 2021-05-04 magic_paper.tex
Tidying of Outlook. More tidying to do.

\section{Free energy dependent bounds for magic state distillation protocols}
\label{sec:stab}

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \nick{cite David?}

@@

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality \ddd{[cite]} for approximating the free energy of a system, to obtain bounds in a similar form.

\subsection{Variation of bounds with physical parameters and constraints from complete positivity}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

\ddd{[To be tidied+blended]} We also note that one can obtain another set of general distillation bounds from the monotonicity of discrimination measures under quantum channels. For example, the $L_1$ norm is contractive under quantum channels so that 
\begin{equation}
|| \rho - \sigma||_1 \ge ||\E(\rho) - \E(\sigma)||_1,
\end{equation}
for any quantum channel $\E$. Therefore, beyond the constraint that the magic protocol corresponds to a stochastic map in the Wigner distribution, we also have constraints coming from complete-positivity and trace-preservation. For example, is $D(\rho,\sigma)$ is a contractive measure on pairs of quantum states that is also additive, in the sense that $D(\rho^{\otimes n}, \sigma^{\otimes n}) = n D(\rho, \sigma)$, then it is readily seen that within the context of the above theorem we also have,
\begin{equation}
R \le \frac{D(\rho_S(\epsilon),\tau)}{D(\rho_S(\epsilon'), \tau')},
\end{equation}
provided the denominator does not vanish. In particular we can choose the Renyi-divergences $D_\alpha(\rho ||\tau)$....or more concretely the von Neumann divergence $S(\rho ||\tau) = \tr[ \rho \log \rho - \rho \log \tau]$. However, if $\tau$ is taken to be a Gibbs state then $S(\rho ||\tau) = \beta (F[\rho] - F[\tau])$, namely the difference in free energies of the two states. Here we define
\begin{equation}
F[\rho] := \tr[H \rho] - \beta^{-1} S(\rho).
\end{equation}
  Therefore, simply from the fact that the protocol defines a channel, we obtain the bound
\begin{equation}
R \le \frac{F[\rho_S(\epsilon)] - F[\tau]}{F[\rho_S(\epsilon')] - F[\tau']},
\end{equation}
and so we can compare this with the above bounds \ddd{[Some numerics on this to compare?]}

@@

\nick{FIX} The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}

@@

\ddd{[Initial text. To be edited.]}
We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Given this, one might wonder if this approach could also provide \emph{lower bounds} for distillation protocols. Now, in contrast to the upper bounds, it is necessary to include the symplectic constraints on the phase space into the majorization relations.

  For example, the Clifford group action on a quantum system corresponds to the action of the affine symplectic group on the discrete phase space. Therefore, any convex mixture of Clifford unitaries will correpond to a convex mixture of these group actions. From the Hardy-Littlewood theorem, we know that traditional majorization is obtained from convex mixtures of arbitrary permutations. However this has been generalised to convex mixtures of elements of a group $G$ that acts on vectors. This $G$--majorization has been studied in the classical literature and a range of results are known about it. For example, if $G$ is a finite reflection group then the $G$--majorization pre-order is described by a finite list of conditions, just as happens for the full majorization ordering. One route is therefore to consider reflection sub-groups of the affine symplectic group, for which the majorization relation reduces to a finite set of conditions. This would therefore provide concrete lower bounds on distillation rates.

  The topic of $G$--majorization has been extensively studied, but to our knowledge there hasn't been work on relative $G$--majorization. This would correspond to transformations that are \emph{not} unital, but still respect the symplectic structure. Physically, this regime would correspond to a form of thermo-majorization obtained from looking at the action of the Clifford group at a micro-canonical level, and then reducing to a small subsystem. While this seems like a painful thing to consider, there is motivation for this beyond the aim of constructing lower bounds for magic protocols: in the case of statistical mechanics on a phase space this is precisely the situation, albeit in the continuum limit. Statistical mechanics of actual systems obey Hamiltonian dynamics and so automatically respect a symplectic form. Therefore the pre-order of statistical mechanical states with respect to phase space dynamics preserving a Gibbs state must correspond to a symplectic thermo-majorization condition. Of course, technical features arise in the continuum limit when considering distributions on an unbounded phase space, but this could be remedied by either considering a compact phase space (e.g. for a particle on a ring) or by first studying the discrete case.

In this work we have looked at magic through the lens of thermodynamics. Specifically, we find that it can be viewed as a strictly non-classical form of non-equilibrium free energy.

Our work also raises a fundamental question that is also of interest in statistical mechanics. Specifically: what are the majorization conditions that respect the symplectic structure of phase space?

>>>>

\newpage
\section{Free energy dependent bounds for magic state distillation protocols}
\label{sec:stab}

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \ddd{[cite: look up papers by Marcus Mueller. He has about 3 or 4 on this. Also Mark Wilde has a paper on it]}

@@

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality \ddd{[cite]} for approximating the free energy of a system, to obtain bounds in a similar form.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\subsection{Variation of bounds with physical parameters}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

@@

\ddd{[Need some numerics on varying the $F'$ case]}
\nick{FIX} The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.

@@

\ddd{[Initial text. To be edited.]}
We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Our bounds only exploited the very simplest aspects of the Lorenz curves of the quasi-distributions, and so it would be of interest to sharpen these bounds and exploit more of the structure in the asymptotic limit. This raises interesting questions that do no arise for probability distributions, because we no longer have a notion of typicality and the central limit theorem does not apply. That said, for special states such as the Strange state the asymptotic behaviour is relatively simple and so one could study this and obtain a better handle on the $n\rightarrow \infty$ behaviour in a way. In appendix (\ddd{BLAH} we give some basic relations on this that could be explored further. \ddd{[where did the remainder term bounds go? People could potentially build on this]}

One might wonder if this approach could also provide \emph{lower bounds} for distillation protocols. Now, in contrast to the upper bounds, it is necessary to include the symplectic constraints on the phase space into the majorization relations.

  For example, the Clifford group action on a quantum system corresponds to the action of the affine symplectic group \ddd{[cite Gross discrete Hudson paper, and some maths source on symplectic group]} on the discrete phase space. Therefore, any convex mixture of Clifford unitaries will correpond to a convex mixture of these group actions. From the Hardy-Littlewood theorem \ddd{[cite]}, we know that traditional majorization is obtained from convex mixtures of arbitrary permutations. However this has been generalised to convex mixtures of elements of a group $G$ that acts on vectors. This $G$--majorization has been studied in the classical literature and a range of results are known about it \ddd{[cite papers]}. For example, if $G$ is a finite reflection group then the $G$--majorization pre-order is described by a finite list of conditions\ddd{[cite]}, just as happens for the full majorization ordering. One route is therefore to consider reflection sub-groups of the affine symplectic group, for which the majorization relation reduces to a finite set of conditions. This would therefore provide concrete lower bounds on distillation rates.

  The topic of $G$--majorization has been extensively studied, but to our knowledge there hasn't been work on relative $G$--majorization. This would correspond to transformations that are \emph{not} unital, but still respect the symplectic structure. Physically, this regime would correspond to a form of thermo-majorization obtained from looking at the action of the Clifford group at a micro-canonical level, and then reducing to a small subsystem \ddd{[cite stat-mech book, cite Matteo review]}. While this seems like a painful thing to consider, there is motivation for this beyond the aim of constructing lower bounds for magic protocols: in the case of statistical mechanics on a phase space this is precisely the situation, albeit in the continuum limit. Statistical mechanics of actual systems obey Hamiltonian dynamics and so automatically respect a symplectic form \ddd{[cite a stat-mech book]}. Therefore the pre-order of statistical mechanical states with respect to phase space dynamics preserving a Gibbs state must correspond to a symplectic thermo-majorization condition. Of course, technical features arise in the continuum limit when considering distributions on an unbounded phase space, but this could be remedied by either considering a compact phase space (e.g. for a particle on a ring) or by first studying the discrete case. Such scenarios also arise in the Quantum Hall Effect \ddd{[cite]} and so this is another regime in which these techniques might be of use.

Finally, it would be of interest to consider the possibility of applying these techniques to other scenarios in which one wishes to distinguish classical from non-classical behaviour based on quasi-probability representations \ddd{[cite quantum optics reference, cite Ferrie review on quasi-reps]}
